IML Term project

### 2022-12-07 main_markdown.rmd
combining files

```{r load data}
npf <- read.csv("npf_train.csv")
```

@@

# Dataset names:
# npf          - "Original" data: has class2 with values 0,1 as numeric values, 
#                   columns removed: class4, partlybad, date (moved to rowname)
# npf.fact     - Otherwise the same as npf, but class2 is a factor with levels 0 and 1
# npf.168      - Processed data: Parameters that have different measurement heights have been reduced to height 16.8m
#                   columns removed: Parameters with height that's not 16.8m
# npf.168.fact -  Otherwise the same as npf.168, but class2 is a factor
# npfsp        - Reduced set of explanatory variables, received through running VIF analysis and repeating variable with highest VIF until no variable has a VIF
#                 higher than 10
# npfsp.fact   - Otherwise the same as npfsp but class2 is a factor

# Create class2, 0 = nonevent, 1 = event
npf$class2 <- ifelse(npf$class4 == "nonevent", 0, 1)

# Add date as row name
rownames(npf) <- npf[, "date"]

# Select global and 16.8 m high values
npf.168 <- data.frame(npf[, (5:6)], npf[, (13:16)], npf[, (27:30)], 
                  npf[, (41:42)], npf[, (53:54)], npf[, (63:72)],
                  npf[, (83:88)], npf[, (99:105)])

# Remove columns id, date, class4, partlybad from npf
npf <- npf[, -(1:4)]

@@

### vif
```{r vif}
testmod<-glm(class2~., npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std-RGlob.std-RGlob.mean, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std-RGlob.std-RGlob.mean-CS.mean-CS.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))
testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std-RGlob.std-RGlob.mean-CS.mean-CS.std-NOx168.mean-NOx168.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

#Thus, we come up to the following collection of variables
coefficients(testmod)
colnames(npf)
npfsp<-npf.168[,c(1:2,5:6,9:10,13:16,19:20,23:24,27:30,37)]
colnames(npfsp)
npfsp.fact<-npfsp
npfsp.fact[,19]<-as.factor(npfsp.fact[,19)
```

@@

```{r correlation at different heights, message=FALSE}
# Korrelaatiomatriisi
cm <- cor(npf[, endsWith(colnames(npf), ".mean")])
# Remove .mean from column and row names.
colnames(cm) <- rownames(cm) <- sapply(colnames(cm), function(s) gsub(".mean", "", s))
# Muuttujakohtaiset korrelaatiot (kun enemmän kuin yksi muuttuja, jolla sama alkuosa)
kable(cm[startsWith(colnames(cm), "CO2"), startsWith(colnames(cm), "CO2")],
      caption = "Correlation (CO2)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "H2O"), startsWith(colnames(cm), "H2O")],
      caption = "Correlation (H20)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "NO") & !startsWith(colnames(cm), "NOx"), startsWith(colnames(cm), "NO") & !startsWith(colnames(cm), "NOx")],
      caption = "Correlation (NO)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "NOx"), startsWith(colnames(cm), "NOx")],
      caption = "Correlation (NOx)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "O3"), startsWith(colnames(cm), "O3")],
      caption = "Correlation (O3)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "RHIRGA"), startsWith(colnames(cm), "RHIRGA")],
      caption = "Correlation (RHIRGA)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "T"), startsWith(colnames(cm), "T")],
      caption = "Correlation (T)") %>%
  kable_styling(latex_options = "HOLD_position")
```

>>>>

```{r load data}
npfOrig <- read.csv("npf_train.csv")
```

@@

# Dataset names:
# npfOrig          - Original data with has class2 as numeric values 0, 1
# npf          - "Original" data: has class2 with values 0,1 as numeric values, 
#                   columns removed: class4, partlybad, date (moved to rowname)
# npf.fact     - Otherwise the same as npf, but class2 is a factor with levels 0 and 1
# npf.168      - Processed data: Parameters that have different measurement heights have been reduced to height 16.8m
#                   columns removed: Parameters with height that's not 16.8m
# npf.168.fact -  Otherwise the same as npf.168, but class2 is a factor
# npfsp        - Reduced set of explanatory variables, received through running VIF analysis and repeating variable with highest VIF until no variable has a VIF
#                 higher than 10
# npfsp.fact   - Otherwise the same as npfsp but class2 is a factor

# Create class2, 0 = nonevent, 1 = event
npfOrig$class2 <- ifelse(npfOrig$class4 == "nonevent", 0, 1)

# Add date as row name
rownames(npfOrig) <- npfOrig[, "date"]

# Select global and 16.8 m high values
npf.168 <- data.frame(npfOrig[, (5:6)], npfOrig[, (13:16)], npfOrig[, (27:30)], 
                  npfOrig[, (41:42)], npfOrig[, (53:54)], npfOrig[, (63:72)],
                  npfOrig[, (83:88)], npfOrig[, (99:105)])

# Remove columns id, date, class4, partlybad from npf
npf <- npfOrig[, -(1:4)]

@@

### vif
```{r vif}
library(car)

testmod<-glm(class2~., npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

testmod<-glm(class2~.-PAR.mean-PAR.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std-RGlob.std-RGlob.mean, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std-RGlob.std-RGlob.mean-CS.mean-CS.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

testmod<-glm(class2~.-PAR.mean-PAR.std-UV_A.std-UV_A.mean-Glob.std-Glob.mean-UV_B.std-UV_B.mean-RPAR.mean-RPAR.std-NET.mean-NET.std-RGlob.std-RGlob.mean-CS.mean-CS.std-NOx168.mean-NOx168.std, npf.168,family=binomial(link=logit))
inflfact<-vif(testmod)
max(inflfact)
which(inflfact==max(inflfact))

#Thus, we come up to the following collection of variables
coefficients(testmod)
colnames(npf)
npfsp<-npf.168[,c(1:2,5:6,9:10,13:16,19:20,23:24,27:30,37)]
colnames(npfsp)
npfsp.fact<-npfsp
npfsp.fact[,19]<-as.factor(npfsp.fact[,19])
```

@@

```{r correlation at different heights, message=FALSE}
# Correlation matrix
cm <- cor(npf[, endsWith(colnames(npf), ".mean")])
# Remove .mean from column and row names.
colnames(cm) <- rownames(cm) <- sapply(colnames(cm), function(s) gsub(".mean", "", s))
# Muuttujakohtaiset korrelaatiot (kun enemmän kuin yksi muuttuja, jolla sama alkuosa)
kable(cm[startsWith(colnames(cm), "CO2"), startsWith(colnames(cm), "CO2")],
      caption = "Correlation (CO2)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "H2O"), startsWith(colnames(cm), "H2O")],
      caption = "Correlation (H20)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "NO") & !startsWith(colnames(cm), "NOx"), startsWith(colnames(cm), "NO") & !startsWith(colnames(cm), "NOx")],
      caption = "Correlation (NO)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "NOx"), startsWith(colnames(cm), "NOx")],
      caption = "Correlation (NOx)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "O3"), startsWith(colnames(cm), "O3")],
      caption = "Correlation (O3)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "RHIRGA"), startsWith(colnames(cm), "RHIRGA")],
      caption = "Correlation (RHIRGA)") %>%
  kable_styling(latex_options = "HOLD_position")
kable(cm[startsWith(colnames(cm), "T"), startsWith(colnames(cm), "T")],
      caption = "Correlation (T)") %>%
  kable_styling(latex_options = "HOLD_position")
```
